Unrecognized States: A Theory of Self-Determination and Foreign Influence

### 2016-03-08 draft.tex
replace hardcoding of restriction numbering with hyperref coding

\begin{enumerate}
\item \textit{For both players $g$ and $s$, $Q_i^0 \geq L_i^0$: remaining in the status quo is better than ceding at the beginning of the game.}

\item \textit{For both players $g$ and $s$, $\frac {Q_i^0}{1-\delta} \geq  -\zeta_i+\frac{W_i^0(p_i) + L_i^0(1-p_i)}{1-\delta}$: the expected outcome under war is worse than the status quo at the beginning of the game.}

\item \textit{$\frac{\alpha}{1-\de} \geq \frac{\beta}{1-\de} + \mu$: reunification is more important for the patron to avoid than for the international community to achieve.}

\item  \textit{$\nu\cdot p_s \geq \lambda \cdot p_s + \mu + \beta$: recognition of the secessionist state is more important for the international community to avoid than for the patron to achieve.}

\item  \textit{$B_p^1 \geq\frac{\beta}{1-\delta} - \left(Q_s^0 -\mu - L_s^0 \right)$: the patron can afford to deter player $c$ from inducing reunification at the beginning of the game.}

\item \textit{$B_p^t \geq \mu, \ \forall t>1$: the patron can afford to pay to maintain the status quo.}\footnote{Depending on parameters, one might think that Restriction (5) is more likely to be binding than Restriction (6). If there is great variance in budget between periods for the patron, such as a greatly reduced budget in some period $t> 1$ compared to period $1$, (6) could be the binding constraint.}

@@

Because the patron's preferences are aligned with the secessionists and against the government, it never invests in $W_g$ or $L_s$. The only reason the patron would invest in $Q_g$ is to counter an investment by player $c$ in the government's payoffs from war; since player $c$ will not make such an investment, we can also rule out investments by the patron in $Q_g$. This leaves four state variables in which the patron may invest: $Q_s$, $L_g$ and $W_s$.

@@

\begin{lemma}
When Restriction (3) of Definition 1 holds, the patron's willingness to invest to maintain the status quo is sufficient to deter the international community from intervening to encourage reunification.
\end{lemma}

\noindent Restriction (3) provides a bound on the amount that player $p$ must be \emph{willing} to invest each period in order to prevent player $c$ from contesting the status quo outcome. To complete the equilibrium construction, we must determine the utility maximizing investments by $p$ and conditions to ensure that it is \emph{able} to make those investments.\footnote{Efforts by the patron to create the conditions for the \emph{Recognition} outcome are addressed in Lemma 3 once the equilibrium investments are established.}

Consider period 1. When Restriction (3) holds, the patron will want to invest just enough to create a buffer of $\frac{\beta}{1-\de}$ between $Q_s^1$ and $L_s^1$ so that $c$ will not invest. That is, in equilibrium, $R_p^1 = \frac{\beta}{1-\de} -(Q_s^0 - \mu - L_s^0)$ as long as $p$ can afford to make this investment. Thus we need Restriction (5): $B_p^1 \geq\frac{\beta}{1-\de} - \left(Q_s^0 - \mu - L_s^0 \right)$, where $B_p^1$ is the amount $p$ has available to spend on the conflict in period $1$.

In periods $t > 1$, there are two cases to consider. Either (a) the buffer created in period $t-1$ was precisely the necessary $\frac{\beta}{1-\de}$, or (b) the buffer is larger than $\frac{\beta}{1-\de}$. In case (a), the patron must spend exactly $\mu$ to offset the degradation in the status quo payoffs and re-establish the buffer of $\frac{\beta}{1-\de}$.\footnote{A third case in which the buffer is smaller than $\frac{\beta}{1-\de}$ provides lower welfare to player $p$. Restrictions (3) and (5) ensure that player $p$ can avoid this case.} In case (b), the patron can spend less than $\mu$ in period $t$. However, because in each period $Q_s^t$ degrades by $\mu$, eventually the buffer will be reached and we will be returned to case (a). Again, assuming Restriction (3) holds, the patron will want to make this investment if its budget allows, and so a sufficient condition is that $p$'s budget is at least as large as $\mu$ in every period $t > 1$ (Restriction 6).

@@

\begin{lemma}
When Restriction (4) of Definition 1 holds, the international community's willingness to invest to avoid recognition of the secessionist state is sufficient to deter the patron from investing to achieve recognition.
\end{lemma}

In the Status Quo equilibrium, the patron will invest to maintain an outcome that is not its most preferred, but it will not invest to achieve its most preferred outcome. This behavior may appear counterintuitive, but we frequently observe patron states whose preferred outcomes are recognized independence for the secessionists who  nonetheless contribute resources to sustain a status-quo outcome that is costly to all involved. The patron does not attempt to contribute sufficient resources to force recognition by the home state because doing so would induce offsetting expenditures by the international community to prevent this outcome.\footnote{We do not provide conditions on the budget of player $c$ similar to Restrictions (5) and (6): since the size of the international community relative to any particular country is large, it can be assumed that a budget constraint does not bind.}

@@

\begin{lemma}
Restrictions (3) and (4) of Definition 1 ensure that the international community's willingness to invest to discourage new conflict is sufficient to deter the patron from investing to instigate such fighting.
\end{lemma}

@@

First, note that the restrictions in Definition 1 are sufficient but not necessary. For example, if the status quo initially has a much higher long term payoff than the next best alternative for the secessionists, Restriction (6) need not be met to maintain the status quo in the short run. The inequality \emph{will} bind for some set of periods $t \geq 1$ because the secessionist payoffs from the status quo decrease over time. In cases where Restriction (6) is not met, we can have unrecognized statehood for some time, but it is not a long-run equilibrium outcome.

@@

Restriction (1) combined with equilibrium play implies that  $Q_s^{t-1} > L_s^{t-1}$. Combining this with $R_c^{SQ} =0$, a tighter condition for preventing player $c$ from inducing reunification is $\frac{\beta}{1 -\de} + \mu \leq R_p^t$.

@@

Restriction (1) combined with equilibrium play implies that  $Q_s^{t-1} > L_s^{t-1}$ so that a tighter condition for preventing player $p$ from inducing recognition is $\frac{\nu}{1 -\de} \geq R_p^t$.

Since the patron is willing to invest up to $\frac{\lambda + \beta + \mu}{1 -\de} -\mu$, when $\frac{\nu}{1 -\de} \geq \frac{\lambda + \beta + \mu}{1 -\de} -\mu$, player $c$'s willingness to counter the patron's investments is sufficient to prevent the \emph{recognition} outcome. A slightly weaker version of this inequality is stated as Restriction (4) so that this deterrence dynamic is guaranteed to operate through the war lottery as well, as will be seen in Lemma 4. \hfill $\blacksquare$
\\
\\
\noindent {\bf Proof of Lemma 4} \\
The difference in continuation values from switching from the status quo to taking the war lottery for player $p$ is $\frac{\lambda p_s - \alpha (1-p_s)}{1 -\de} +R_p^{SQ} -R_p^{WAR}$. For player $c$ it is $\frac{-\nu p_s + \beta (1-p_s)}{1 -\de} + R_c^{SQ} -R_c^{WAR}$. As noted above, $R_c^{SQ} =0$. So $c$ is willing to invest up to $\frac{\nu p_s - \beta (1-p_s)}{1 -\de}$ toward preventing fighting.

@@

Restriction (1) combined with equilibrium play implies that  $Q_s^{t-1} \geq \omega_s^{t-1}$ so a tighter condition for preventing player $p$ from inducing fighting is $\frac{\nu p_s - \beta (1-p_s)}{1 -\de} - \mu \geq R_p^t p_s$.

@@

Multiplying Restriction (3) by $-(1-p_s)$, we have $\frac{-\beta \cdot (1-p_s)}{1-\de} -\mu (1-p_s) \geq \frac{-\alpha \cdot (1-p_s)}{1-\de}$. Using this with Restriction (4) multiplied by $\frac{1}{1-\de}$, we have $\frac{\nu p_s - \beta (1-p_s)}{1-\de} -\mu \geq \frac{\lambda p_s - \alpha (1-p_s) + \mu + \beta}{1-\de} -p_s \mu$. Since it is assumed that the war lottery probabilities are non-degenerate, we also have $\frac{\lambda p_s - \alpha (1-p_s) + \mu + \beta}{1-\de} -p_s \mu > p_s \frac{\lambda p_s - \alpha (1-p_s) + \mu + \beta}{1-\de} -p_s \mu$. \hfill $\blacksquare$
\\
\\
{\bf Proof of Proposition 2} \\
Suppose that sanctions are applied by player $c$ in period $t+1$ after a Status Quo equilibrium such as the one supported by the construction in Proposition 1 has been established in period $t$. As the only alteration is a reduction in $Q_s^t$ by the amount of the sanctions, it is immediate that if player $p$ is willing and able to supplement player $s$'s status quo payoffs to counteract sanctions, then player $s$'s incentives to play Status Quo are not disturbed. As in Proposition 1, it is not necessary that the patron make these payments immediately if player $s$'s utility from playing Status Quo is high enough relative to its other options, but the degradation of the status quo payoffs will eventually make these payments necessary in order to sustain a Status Quo equilibrium.

>>>>

\begin{enumerate}
\item \textit{For both players $g$ and $s$, $Q_i^0 \geq L_i^0$: remaining in the status quo is better than ceding at the beginning of the game.}\label{res:1}

\item \textit{For both players $g$ and $s$, $\frac {Q_i^0}{1-\delta} \geq  -\zeta_i+\frac{W_i^0(p_i) + L_i^0(1-p_i)}{1-\delta}$: the expected outcome under war is worse than the status quo at the beginning of the game.}\label{res:2}

\item \textit{$\frac{\alpha}{1-\de} \geq \frac{\beta}{1-\de} + \mu$: reunification is more important for the patron to avoid than for the international community to achieve.}\label{res:3}

\item  \textit{$\nu\cdot p_s \geq \lambda \cdot p_s + \mu + \beta$: recognition of the secessionist state is more important for the international community to avoid than for the patron to achieve.}\label{res:4}

\item  \textit{$B_p^1 \geq\frac{\beta}{1-\delta} - \left(Q_s^0 -\mu - L_s^0 \right)$: the patron can afford to deter player $c$ from inducing reunification at the beginning of the game.}\label{res:5}

\item \textit{$B_p^t \geq \mu, \ \forall t>1$: the patron can afford to pay to maintain the status quo.}\footnote{Depending on parameters, one might think that Restriction (5) is more likely to be binding than Restriction (6). If there is great variance in budget between periods for the patron, such as a greatly reduced budget in some period $t> 1$ compared to period $1$, (6) could be the binding constraint.}\label{res:6}

@@

Because the patron's preferences are aligned with the secessionists and against the government, it never invests in $W_g$ or $L_s$. The only reason the patron would invest in $Q_g$ is to counter an investment by player $c$ in the government's payoffs from war; since player $c$ will not make such an investment, we can also rule out investments by the patron in $Q_g$. This leaves three state variables in which the patron may invest: $Q_s$, $L_g$ and $W_s$.

@@

\begin{lemma}
When Restriction (\ref{res:3}) of Definition 1 holds, the patron's willingness to invest to maintain the status quo is sufficient to deter the international community from intervening to encourage reunification.
\end{lemma}

\noindent Restriction (\ref{res:3}) provides a bound on the amount that player $p$ must be \emph{willing} to invest each period in order to prevent player $c$ from contesting the status quo outcome. To complete the equilibrium construction, we must determine the utility maximizing investments by $p$ and conditions to ensure that it is \emph{able} to make those investments.\footnote{Efforts by the patron to create the conditions for the \emph{Recognition} outcome are addressed in Lemma 3 once the equilibrium investments are established.}

Consider period 1. When Restriction (\ref{res:3}) holds, the patron will want to invest just enough to create a buffer of $\frac{\beta}{1-\de}$ between $Q_s^1$ and $L_s^1$ so that $c$ will not invest. That is, in equilibrium, $R_p^1 = \frac{\beta}{1-\de} -(Q_s^0 - \mu - L_s^0)$ as long as $p$ can afford to make this investment. Thus we need Restriction (\ref{res:5}): $B_p^1 \geq\frac{\beta}{1-\de} - \left(Q_s^0 - \mu - L_s^0 \right)$, where $B_p^1$ is the amount $p$ has available to spend on the conflict in period $1$.

In periods $t > 1$, there are two cases to consider. Either (a) the buffer created in period $t-1$ was precisely the necessary $\frac{\beta}{1-\de}$, or (b) the buffer is larger than $\frac{\beta}{1-\de}$. In case (a), the patron must spend exactly $\mu$ to offset the degradation in the status quo payoffs and re-establish the buffer of $\frac{\beta}{1-\de}$.\footnote{A third case in which the buffer is smaller than $\frac{\beta}{1-\de}$ provides lower welfare to player $p$. Restrictions (\ref{res:3}) and (\ref{res:5}) ensure that player $p$ can avoid this case.} In case (b), the patron can spend less than $\mu$ in period $t$. However, because in each period $Q_s^t$ degrades by $\mu$, eventually the buffer will be reached and we will be returned to case (a). Again, assuming Restriction (\ref{res:3}) holds, the patron will want to make this investment if its budget allows, and so a sufficient condition is that $p$'s budget is at least as large as $\mu$ in every period $t > 1$ (Restriction \ref{res:6}).

@@

\begin{lemma}
When Restriction (\ref{res:4}) of Definition 1 holds, the international community's willingness to invest to avoid recognition of the secessionist state is sufficient to deter the patron from investing to achieve recognition.
\end{lemma}

In the Status Quo equilibrium, the patron will invest to maintain an outcome that is not its most preferred, but it will not invest to achieve its most preferred outcome. This behavior may appear counterintuitive, but we frequently observe patron states whose preferred outcomes are recognized independence for the secessionists who  nonetheless contribute resources to sustain a status-quo outcome that is costly to all involved. The patron does not attempt to contribute sufficient resources to force recognition by the home state because doing so would induce offsetting expenditures by the international community to prevent this outcome.\footnote{We do not provide conditions on the budget of player $c$ similar to Restrictions (\ref{res:5}) and (\ref{res:6}): since the size of the international community relative to any particular country is large, it can be assumed that a budget constraint does not bind.}

@@

\begin{lemma}
Restrictions (\ref{res:3}) and (\ref{res:4}) of Definition 1 ensure that the international community's willingness to invest to discourage new conflict is sufficient to deter the patron from investing to instigate such fighting.
\end{lemma}

@@

First, note that the restrictions in Definition 1 are sufficient but not necessary. For example, if the status quo initially has a much higher long term payoff than the next best alternative for the secessionists, Restriction (\ref{res:6}) need not be met to maintain the status quo in the short run. The inequality \emph{will} bind for some set of periods $t \geq 1$ because the secessionist payoffs from the status quo decrease over time. In cases where Restriction (\ref{res:6}) is not met, we can have unrecognized statehood for some time, but it is not a long-run equilibrium outcome.

@@

Restriction (\ref{res:1}) combined with equilibrium play implies that  $Q_s^{t-1} > L_s^{t-1}$. Combining this with $R_c^{SQ} =0$, a tighter condition for preventing player $c$ from inducing reunification is $\frac{\beta}{1 -\de} + \mu \leq R_p^t$.

@@

Restriction (\ref{res:1}) combined with equilibrium play implies that  $Q_s^{t-1} > L_s^{t-1}$ so that a tighter condition for preventing player $p$ from inducing recognition is $\frac{\nu}{1 -\de} \geq R_p^t$.

Since the patron is willing to invest up to $\frac{\lambda + \beta + \mu}{1 -\de} -\mu$, when $\frac{\nu}{1 -\de} \geq \frac{\lambda + \beta + \mu}{1 -\de} -\mu$, player $c$'s willingness to counter the patron's investments is sufficient to prevent the \emph{recognition} outcome. A slightly weaker version of this inequality is stated as Restriction (\ref{res:4}) so that this deterrence dynamic is guaranteed to operate through the war lottery as well, as will be seen in Lemma 4. \hfill $\blacksquare$
\\
\\
\noindent {\bf Proof of Lemma 4} \\
The difference in continuation values from switching from the status quo to taking the war lottery for player $p$ is $\frac{\lambda p_s - \alpha (1-p_s)}{1 -\de} +R_p^{SQ} -R_p^{WAR}$. For player $c$ it is $\frac{-\nu p_s + \beta (1-p_s)}{1 -\de} + R_c^{SQ} -R_c^{WAR}$. As noted above, $R_c^{SQ} =0$. So $c$ is willing to invest up to $\frac{\nu p_s - \beta (1-p_s)}{1 -\de}$ toward preventing fighting.

@@

Restriction (\ref{res:1}) combined with equilibrium play implies that  $Q_s^{t-1} \geq \omega_s^{t-1}$ so a tighter condition for preventing player $p$ from inducing fighting is $\frac{\nu p_s - \beta (1-p_s)}{1 -\de} - \mu \geq R_p^t p_s$.

@@

Multiplying Restriction (\ref{res:3}) by $-(1-p_s)$, we have $\frac{-\beta \cdot (1-p_s)}{1-\de} -\mu (1-p_s) \geq \frac{-\alpha \cdot (1-p_s)}{1-\de}$. Using this with Restriction (\ref{res:4}) multiplied by $\frac{1}{1-\de}$, we have $\frac{\nu p_s - \beta (1-p_s)}{1-\de} -\mu \geq \frac{\lambda p_s - \alpha (1-p_s) + \mu + \beta}{1-\de} -p_s \mu$. Since it is assumed that the war lottery probabilities are non-degenerate, we also have $\frac{\lambda p_s - \alpha (1-p_s) + \mu + \beta}{1-\de} -p_s \mu > p_s \frac{\lambda p_s - \alpha (1-p_s) + \mu + \beta}{1-\de} -p_s \mu$. \hfill $\blacksquare$
\\
\\
{\bf Proof of Proposition 2} \\
Suppose that sanctions are applied by player $c$ in period $t+1$ after a Status Quo equilibrium such as the one supported by the construction in Proposition 1 has been established in period $t$. As the only alteration is a reduction in $Q_s^t$ by the amount of the sanctions, it is immediate that if player $p$ is willing and able to supplement player $s$'s status quo payoffs to counteract sanctions, then player $s$'s incentives to play Status Quo are not disturbed. As in Proposition 1, it is not necessary that the patron make these payments immediately if player $s$'s utility from playing Status Quo is high enough relative to its other options, but the degradation of the status quo payoffs will eventually make these payments necessary in order to sustain a Status Quo equilibrium.
